---
title: "lat and long"
author: "Richard G. Gardiner"
date: "May 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

## Haversine Formula (as the Crow Flies)

Uses the paper "Virtues of the Haversine", sky and Telescope Volume 68, no. 2, 1984, page 150 by R W. Sinott

formulas:

- dlon = longitude2 - longitude 1
- dlat = lat2 - lat1
- a = (sin(dlat/2))^2 + cos(lat1) * cos(lat2) * sin(dlon/2))^2
- c = 2* atan2(sqrt(a), sqrt(1-a))
- d = R * c

```{r}
library(pracma)
library(tidyverse)
library(readxl)
```

Testing the haversine for two random locations and then comparing with google:
```{r}
albany <- c(29.93601, -77.0889) # location for albany
arizona <- c(33.41772, -111.934) # location for arizona

distance <- haversine(albany, arizona) # how you actually run this

distance * 0.621371 # converts to miles
```

2053.416 miles is the distance between the albany and arizona schools.



```{r}
FRA <- c(50.0379, 8.5622) # frankfurt airport
ORD <- c(41.9742, -87.9073) # O-Haire

dis <- haversine(FRA, ORD) 

dis * 0.621371 # converting to miles
```

4331.437 miles


Initial testing appears to be pretty good. It give me the distance in kilometers, which I can easily change to miles


## Writing my own function

It would appear that the haversine formula is having trouble understand my dataset, so I decided to try to take the formula from online and apply it here.  Each step basically follows the haversine function that is from the `pracma` package.  

```{r}
haversine_r <- function(lon1, lat1, lon2, lat2){
    # convert decimal degrees to radians
    lon1 <- lon1 * pi / 180
    lon2 <- lon2 * pi / 180
    lat1 <- lat1 * pi / 180
    lat2 <- lat2 * pi / 180
    # haversine formula
    dlon <- lon2 - lon1
    dlat <- lat2 - lat1
    a <- sin(dlat / 2) ** 2 + cos(lat1) * cos(lat2) * sin(dlon/2) ** 2
    c <- 2 * atan2(sqrt(a), sqrt(1 - a))
    km <- 6367 * c
    miles <- km * 0.621371
    miles
}
```


Now I want to plug in the coordinates from the Frankfurt and O'Haire example and see how the formula worked
```{r}
haversine_r(8.5622, 50.0379, -87.9073, 41.9742)
```

The initial formula gives me roughly the same distance within a few miles.  My estimate is 3 miles off from the other one.  Within reason given that the total distance is a little more than 4300 miles. That makes me off by `r (3 / 4328.718) * 100` percent.

## Haversine over a DataFrame

Now I am going to read in a test dataset that is a subset of our larger dataset we will be using.
```{r}
test_dataset <- read_excel("haversine example.xlsx") 

str(test_dataset)
```


here I am running my function with and adding the results as a column in my dataframe. 

```{r}
test_of_function <- test_dataset %>%
  mutate(miles = round(haversine_r(test_dataset$destination_long, 
                                   test_dataset$destination_lat,
                                   test_dataset$atl_long, 
                                   test_dataset$atl_lat), 2))
```
I decided to round this to two spots just to make this all easier to read.

```{r}
DT::datatable(test_of_function)
```

