---
title: "Surveys"
author: "Richard G. Gardiner"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Packages

```{r}
library(NHANES)
library(tidyverse)
library(survey)
```


Survey: National Health and Nutrition Examination Survey (NHANES)

Goal: Understand the health of adults and children in US

4 stage design:

1. Stage 0: The U.S. is stratified by geography and proportion of minority populations
2. Stage 1: within strata, counties are randomly selected
3. Stage 2: within counties, city blocks arerandomly selected
4. Stage 3: within city blocks, households randomly selected

```{r}
dim(NHANESraw) # 78 variables with 20,293 individuals
```

Before specifying the survey design, we need to modify the survey weights variable

```{r}
NHANESraw %>%
  summarize(N_hat = sum(WTMEC2YR))
```

Our weight should give a rough idea of the U.S. population, but this only shows us 608 million people, but that is because the weights assumed we would have 4 years of data, when we only have two.  So need to divide the weights by 2:

```{r}
NHANESraw <- NHANESraw %>%
  mutate(WTMEC4YR =  WTMEC2YR/2)
```


### Specifying the Design

Survey design specification is KEY to understanding the actual population.

Let's look at the weights real quick
```{r}
#Create table of average survey weights by race
tab_weights <- NHANESraw %>%
  group_by(Race1) %>%
  summarize(avg_wt = mean(WTMEC4YR))

#Print the table
tab_weights
```


1-3 counties were surveyed within each county (why we need nest = TRUE)
```{r}
distinct(NHANESraw, SDMVPSU) 

NHANES_design <- svydesign(data = NHANESraw, strata = ~SDMVSTRA,
                           id = ~SDMVPSU, nest = TRUE,
                           weights = ~WTMEC4YR)

summary(NHANES_design)

# Finding the number of clusters (clusers are nested inside the strata)
# Number of clusters
NHANESraw %>%
  summarize(n_clusters = n_distinct(SDMVSTRA, SDMVPSU))
```

### Summarizing Categorical Variables

The NHANES variable, Depressed, gives the self-reported frequency in which a participant felt depressed. It is only reported for participants aged 18 years or older. Let's explore the distribution of depression in the US.  Let's look at some of the data:

```{r}
# Determine the levels of Depressed
levels(NHANESraw$Depressed)

# Construct a frequency table of Depressed
tab_w <- svytable(~Depressed, design = NHANES_design)

# Determine class of tab_w
class(tab_w)

# Display tab_w
tab_w
```

Now let's graph it:

```{r}
# Add proportions to table
tab_w <- tab_w %>%
  as.data.frame() %>%
  mutate(Prop = Freq/sum(Freq))

# Create a barplot
ggplot(data = tab_w,
       mapping = aes(x = Depressed, y = Prop)) + 
  geom_col()
```

### Exploring Two Categorical Variables

Below we are creating survey weighted contingency tables:
```{r}
race_diabetes <- svytable(~ Race1 + Diabetes,
         design = NHANES_design)

# conditional proportion
race_diabetes %>%
  as.data.frame() %>%
  ggplot(aes(x = Race1, y = Freq, fill = Diabetes)) +
  geom_col(position = "fill") +
  coord_flip()
```

Now let's formally


```{r}
race_diab_cond <- race_diabetes %>%
  as.data.frame() %>%
  group_by(Race1) %>%
  mutate(n_Race = sum(Freq),
         Prop_Depressed = Freq/n_Race) # creates the proportion of diabetes for each race

race_diab_cond
```

### Summarizing with svytotal()

We canalso estimate counts with svtotal().  For each combination of the two variables, we get an estimate of the total and the standard error.

```{r}
tab_totals <- svytotal(x = ~interaction(Race1, Diabetes),
                       design = NHANES_design,
                       na.rm = TRUE)

tab_totals
```

Now use svymean() to find estimated means for the two:

```{r}
tab_means <- svymean(x = ~interaction(Race1, Diabetes),
                     design = NHANES_design,
                     na.rm = TRUE)

tab_means
```

### making an inference for categorical variables

The code below does a Chi squared test for the relationship between Race and Diabetes.
```{r}
svychisq(~Race1 + Diabetes, design = NHANES_design, statistic = "Chisq")
```



## Summarizing Quantitative Data

The raw data isn't very insightful:
```{r}
NHANESraw %>%
  filter(Age >= 12) %>%
  select(DaysPhysHlthBad)
```

But we can use different functions to get means, medians, survey standard errors.  A particularly helpful function is: svyby() which allows us to compute summaries for different subgroups


How many hours of sleep does the average person get?  What about broken down by gender?
```{r}
# Compute the survey-weighted mean
svymean(x = ~SleepHrsNight, 
       design = NHANES_design,
       na.rm = TRUE)

# Compute the survey-weighted mean by Gender
svyby(formula = ~SleepHrsNight, 
    by = ~Gender, 
    design = NHANES_design, 
    FUN = svymean, 
    na.rm = TRUE, 
    keep.names = FALSE)
```

Now compute the median and different quantile values:

```{r}
# Compute the survey-weighted quantiles
svyquantile(x = ~SleepHrsNight, 
            design = NHANES_design, 
            na.rm = TRUE, 
            quantiles = c(0.01, 0.25, 0.5, 0.75, .99))

# Compute the survey-weighted quantiles by Gender
svyby(formula = ~SleepHrsNight, 
      by = ~Gender, 
      design = NHANES_design, 
      FUN = svyquantile, 
      na.rm = TRUE, 
      quantiles = 0.5, 
      keep.rows = FALSE, 
      keep.var = FALSE)
```



### Visualizing quantitative data:

```{r}
# Compute the survey-weighted mean by Gender
out <- svyby(formula = ~SleepHrsNight, 
             by = ~Gender, 
             design = NHANES_design, 
             FUN = svymean, 
             na.rm = TRUE, 
             keep.names = FALSE)
             
# Construct a bar plot of average sleep by gender
ggplot(data = out, mapping = aes(x = Gender, y = SleepHrsNight)) +
  geom_col() + 
  labs(y = "Average Nightly Sleep")
```

Adding error bars:

```{r}
# Add lower and upper columns to out
out_col <- mutate(out, 
                  lower = SleepHrsNight - 2*se, 
                  upper = SleepHrsNight + 2*se)

# Construct a bar plot of average sleep by gender with error bars
ggplot(data = out_col, 
       mapping = aes(x = Gender, y = SleepHrsNight, 
                     ymin = lower, ymax = upper)) +
  geom_col(fill = "gold") +
  labs(y = "Average Nightly Sleep") +
  geom_errorbar(width = 0.7)
```
If we want to understand the distribution of nightly sleep for Americans, we should construct a histogram. Let's do that now and see the impact of changing the width of the bins.

```{r}
# Create a histogram with a set binwidth
ggplot(data = NHANESraw,
       mapping = aes(SleepHrsNight, weight = WTMEC4YR)) + # note the weight option
  geom_histogram(binwidth = 1,
                 color = "white") +
  labs(x = "Hours of Sleep")
```


Survey weighted density plots:

```{r}
# Density plot of sleep faceted by gender
NHANESraw %>%
    filter(!is.na(SleepHrsNight), !is.na(Gender)) %>%
    group_by(Gender) %>%
    mutate(WTMEC4YR_std = WTMEC4YR/sum(WTMEC4YR)) %>%
    ggplot(mapping = aes(x = SleepHrsNight, weight = WTMEC4YR_std)) + 
        geom_density(bw = 0.6,  fill = "gold") +
        labs(x = "Hours of Sleep") + 
        facet_wrap(~Gender, labeller = "label_both")
```


### Interence for quantitive data


Let's see if there is a difference of sleep hours per night and gender
```{r}
# Run a survey-weighted t-test
svyttest(formula = SleepHrsNight ~ Gender,
       design = NHANES_design)
```



Now let's go through the whole thing:

```{r}
# Find means of total cholesterol by whether or not active 
out <- svyby(formula = ~TotChol,
           by = ~PhysActive, 
           design = NHANES_design,
           FUN = svymean, 
           na.rm = TRUE, 
           keep.names = FALSE)

# Construct a bar plot of means of total cholesterol by whether or not active 
ggplot(data = out,
       mapping = aes(y = TotChol, x = PhysActive)) +
  geom_col()

# Run t test for difference in means of total cholesterol by whether or not active
svyttest(formula = TotChol ~ PhysActive,
    design = NHANES_design)
```

Visualization with scatterplots:

Let's explore the relationship between Height and Weight of Americans. We will focus on 20 year olds to minimize the effects of overplotting. We want to compare the look of the standard scatter plot to a bubble plot.

```{r}
# Create dataset with only 20 year olds
NHANES20 <- filter(NHANESraw,
                Age == 20)

# Construct bubble plot
ggplot(data = NHANES20, 
       mapping = aes(x = Height, y = Weight, size = WTMEC4YR)) + 
    geom_point(alpha = 0.3) + 
    guides(size = FALSE)
```

This time with color:
```{r}
# Construct a scatter plot
ggplot(data = NHANES20,
       mapping = aes(x = Height, y = Weight, color = WTMEC4YR)) + 
    geom_point() + 
    guides(color = FALSE)
```

This time with different alpha levels:

```{r}
# Construct a scatter plot
ggplot(data = NHANES20,
       mapping = aes(x = Height, y = Weight, alpha = WTMEC4YR)) + 
    geom_point() + 
    guides(alpha = FALSE)
```

```{r}
# Add gender to plot
ggplot(data = NHANES20,
       mapping = aes(x = Height, y = Weight, color = WTMEC4YR, color = Gender)) + 
    geom_point(alpha = 0.3) + 
    guides(size = FALSE)
```


```{r}
# Add gender to plot
ggplot(data = NHANES20,
       mapping = aes(x = Height, y = Weight, alpha = WTMEC4YR, color = Gender)) + 
    geom_point(alpha = 0.3) + 
    guides(alpha = FALSE)
```


Visualizing Trends


```{r}
# Add gender to plot
ggplot(data = NHANES20,
       mapping = aes(x = Height, y = Weight, alpha = WTMEC4YR, color = Gender)) + 
    geom_point(alpha = 0.3) + 
    guides(alpha = FALSE)
```

```{r}
# Bubble plot with linear of best fit
ggplot(data = NHANESraw, mapping = aes(x = Height, y = Weight, size = WTMEC4YR)) + 
  geom_point(alpha = 0.1) + 
  guides(size = FALSE) + 
  geom_smooth(method = "lm", se = FALSE, mapping = aes(weight = WTMEC4YR))
```

```{r}
# Add quadratic curve and cubic curve
ggplot(data = NHANESraw, mapping = aes(x = Height, y = Weight, size = WTMEC4YR)) + 
  geom_point(alpha = 0.1) + 
  guides(size = FALSE) + 
  geom_smooth(method = "lm", se = FALSE, mapping = aes(weight = WTMEC4YR)) +
  geom_smooth(method = "lm", se = FALSE, mapping = aes(weight = WTMEC4YR), formula = y ~ poly(x, 2), color = "orange") +
  geom_smooth(method = "lm", se = FALSE, mapping = aes(weight = WTMEC4YR), formula = y ~ poly(x, 3), color = "red")
```

Now lows do the bubble chart colored by gender for 20 year olds. Let's see what happens when we don't account for survey weights and then one that does:

```{r}
# Add non-survey-weighted trend lines to bubble plot
ggplot(data = NHANES20, mapping = aes(x = Height, y = Weight, size = WTMEC4YR, color = Gender)) + 
  geom_point(alpha = 0.1) + 
  guides(size = FALSE) + 
  geom_smooth(method = "lm", se = FALSE, linetype = 2)

# Add survey-weighted trend lines
ggplot(data = NHANES20, mapping = aes(x = Height, y = Weight, size = WTMEC4YR, color = Gender)) + 
  geom_point(alpha = 0.1) + 
  guides(size = FALSE) + 
  geom_smooth(method = "lm", se = FALSE, linetype = 2) + 
  geom_smooth(method = "lm", se = FALSE, mapping = aes(weight = WTMEC4YR))
```




# Modeling Survey Data

Here is how you will use survey to incorporate to weights when you do a regression.
```{r}
# Subset survey design object to only include 20 year olds
NHANES20_design <- subset(NHANES_design, Age == 20)

# Build a linear regression model
mod <- svyglm(Weight ~ Height, design = NHANES20_design)

# Print summary of the model
summary(mod)
```


## More complex modeling (multiple regression)

```{r}
# Build a linear regression model same slope
mod1 <- svyglm(Weight ~ Height + Gender, design = NHANES20_design)

# Print summary of the same slope model
summary(mod1)

# Build a linear regression model different slopes
mod2 <- svyglm(Weight ~ Height * Gender, design = NHANES20_design)

# Print summary of the different slopes model
summary(mod2)
```


# From start to finish:

```{r}
# Plot BPDiaAve and BPSysAve by Diabetes and include trend lines
drop_na(NHANESraw, Diabetes) %>%
ggplot(mapping = aes(x = BPDiaAve, y = BPSysAve, size = WTMEC4YR, color = Diabetes)) + 
    geom_point(alpha = 0.2) + 
    guides(size = FALSE) + 
    geom_smooth(method = "lm", se = FALSE, mapping = aes(weight = WTMEC4YR))

# Build simple linear regression model
mod1 <- svyglm(BPSysAve ~ BPDiaAve, design = NHANES_design)

# Build model with different slopes
mod2 <- svyglm(BPSysAve ~ BPDiaAve * Diabetes, design = NHANES_design)

summary(mod1)
summary(mod2)
```

